---
title: "Recreating Netflix's quantile bootstrapping technique in R"
output:
  html_document
---

This notebook is a full code walkthrough of me Recreating Netflix's quantile bootstrapping technique in R. The details of this analysis are published in [this article on medium](TBC).

# General setup
```{r knit_setup}
knitr::opts_chunk$set(
  comment = "#>",
  fig.path = "temp/"
)
```
```{r setup_packages}
# Will be making heavy use of the tidyverse
library(tidyverse)
```
```{r setup_plots}
# Set default theme as minimal
theme_set(theme_minimal())
```
```{r}
# Set a seed to make the analysis reproducible
set.seed(20200511)
```


# Prep

## Simulating data

```{r}
x <- rchisq(n = 200, df = 8)
```
```{r simulated_data_histogram}
tibble(x) %>% 
  ggplot(aes(x)) +
    geom_histogram(bins = 50)
```

## Getting Quantiles

```{r}
quantile_df <- tibble(percentile = seq(0, 1, by = .2)) %>% 
  mutate(value = quantile(x, percentile))
quantile_df
```

# Checkin: we can do a bit already

```{r}
# Simulate data sets. For treatment, we'll bind two distributions skewed either
# side of production
x_production <- rchisq(200, 8)
x_treatment  <- c(rchisq(100, 12), rchisq(100, 4))
```
```{r}
group_quantiles_df <- tibble(percentile = seq(0, 1, by = .2)) %>% 
  mutate(production = quantile(x_production, percentile),
         treatment  = quantile(x_treatment, percentile))
group_quantiles_df
```

```{r quantile_comparison_absolute}
group_quantiles_df %>% 
  pivot_longer(-percentile, names_to = "group") %>% 
  ggplot(aes(percentile, value)) +
    geom_line(aes(group = group, color = group),
                size = 2) +
    labs(x = "Percentile", y = "Value", color = NULL) +
    ggtitle("Quantile values for two data sets") +
    scale_color_manual(values = c("black", "purple")) +
    scale_x_continuous(lim = c(0, 1), labels = scales::percent_format()) +
    theme(legend.text = element_text(size = 14),
          legend.position = "top")
```

```{r quantile_comparison_relative}
group_quantiles_df %>% 
  mutate(treatment = treatment - production,
         production = 0) %>% 
  pivot_longer(-percentile, names_to = "group") %>% 
  ggplot(aes(percentile, value)) +
    geom_line(aes(group = group, color = group),
              size = 2) +
    labs(x = "Percentile", y = "Value", color = NULL) +
    ggtitle("Difference to production data at various quantiles") +
    scale_color_manual(values = c("black", "purple")) +
    scale_x_continuous(lim = c(0, 1), labels = scales::percent_format()) +
    theme(legend.text = element_text(size = 14),
          legend.position = "top")
```

# Bootstrapping quantile differences

```{r}
x_bootstrap <- sample(x, replace = TRUE)
```
```{r single_bootstrap_comparison}
tibble(x, x_bootstrap) %>% 
  pivot_longer(c(x, x_bootstrap)) %>% 
  ggplot(aes(value, fill = name)) +
    geom_histogram(bins = 50, alpha = .7, position = "identity") +
    labs(fill = NULL) +
    ggtitle("Distribution of x and a bootstrapped sample of it") +
    scale_fill_manual(values = c("black", "green")) +
    theme(legend.text = element_text(size = 14),
          legend.position = "top")
```

```{r}
n_bootstraps <- 100
quantile_df_template <- tibble(percentile = seq(0, 1, by = .2))

bootstrapped_quantiles <- map_df(seq_len(n_bootstraps), function(i) {
  bootstrapped_production <- sample(x_production, replace = TRUE)
  bootstrapped_treatment  <- sample(x_treatment, replace = TRUE)
  
  quantile_df_template %>%
    mutate(bootstrap = i,
           production = quantile(bootstrapped_production, percentile),
           treatment  = quantile(bootstrapped_treatment, percentile))
})
```
```{r}
bootstrapped_quantiles
```

```{r boostrapped_versions_of_data}
bootstrapped_quantiles %>% 
  ggplot(aes(percentile, production)) +
    geom_line(data = group_quantiles_df, size = 3) +
    #geom_line(aes(group = bootstrap), alpha = .1, color = "green") +
    #geom_violin(aes(group = percentile)) +
    stat_summary(fun = mean, fun.min = min, fun.max = max, colour = "red", size = .5) +
    labs(x = "Percentile", y = "Value", color = NULL) +
    ggtitle("Production data (black) and range of bootstrapped versions of it (red)") +
    scale_color_manual(values = c("black", "purple")) +
    scale_x_continuous(lim = c(0, 1), labels = scales::percent_format()) +
    theme(legend.text = element_text(size = 14),
          legend.position = "top")
```

```{r simple_range_overlay}
group_quantiles_df %>% 
  mutate(treatment = treatment - production,
         production = 0) %>% 
  pivot_longer(-percentile, names_to = "group") %>% 
  ggplot(aes(percentile, value)) +
    geom_line(aes(group = group, color = group),
              size = 2) +
    stat_summary(fun = mean, fun.min = min, fun.max = max, colour = "red", size = .5,
                 data = mutate(bootstrapped_quantiles, treatment = treatment - production),
                 aes(y = treatment)) +
    labs(x = "Percentile", y = "Value", color = NULL) +
    ggtitle("Difference to production data with bootstrap ranges") +
    scale_color_manual(values = c("black", "purple")) +
    scale_x_continuous(lim = c(0, 1), labels = scales::percent_format()) +
    theme(legend.text = element_text(size = 14),
          legend.position = "top")
```

# Bootstrapping Confidence Intervals

```{r}
bootstrapped_cis <- bootstrapped_quantiles %>% 
  mutate(q_dif = treatment - production) %>% 
  group_by(percentile) %>% 
  summarise(lower_bound = quantile(q_dif, .025),
            upper_bound = quantile(q_dif, .975))

bootstrapped_cis
```


```{r ci_overlay}
group_quantiles_df %>% 
  mutate(treatment = treatment - production,
         production = 0) %>% 
  pivot_longer(-percentile, names_to = "group") %>% 
  ggplot(aes(percentile)) +
    geom_ribbon(data = bootstrapped_cis,
                aes(ymin = lower_bound, ymax = upper_bound),
                fill = "purple", alpha = .2) +
    geom_line(aes(y = value, group = group, color = group),
              size = 2) +
    labs(x = "Percentile", y = "Value", color = NULL) +
    ggtitle("Difference to production data with bootstrapped confidence intervals") +
    scale_color_manual(values = c("black", "purple")) +
    scale_x_continuous(lim = c(0, 1), labels = scales::percent_format()) +
    theme(legend.text = element_text(size = 14),
          legend.position = "top")
```

# Bump up the numbers

```{r}
n_units <- 1000
quantile_seq <- seq(0, 1, by = .01)
n_bootstraps <- 1000

x_production <- rchisq(n_units, 8)
x_treatment  <- c(rchisq(n_units/2, 12), rchisq(n_units/2, 4))

group_quantiles_df <- tibble(percentile = quantile_seq) %>% 
  mutate(production = quantile(x_production, percentile),
         treatment  = quantile(x_treatment, percentile))

quantile_df_template <- tibble(percentile = quantile_seq)

bootstrapped_quantiles <- map_df(seq_len(n_bootstraps), function(i) {
  bootstrapped_production <- sample(x_production, replace = TRUE)
  bootstrapped_treatment  <- sample(x_treatment, replace = TRUE)
  
  quantile_df_template %>%
    mutate(bootstrap = i,
           production = quantile(bootstrapped_production, percentile),
           treatment  = quantile(bootstrapped_treatment, percentile))
})

bootstrapped_cis <- bootstrapped_quantiles %>% 
  mutate(q_dif = treatment - production) %>% 
  group_by(percentile) %>% 
  summarise(lower_bound = quantile(q_dif, .025),
            upper_bound = quantile(q_dif, .975))
```

```{r ci_overlay_nup}
group_quantiles_df %>% 
  mutate(treatment = treatment - production,
         production = 0) %>% 
  pivot_longer(-percentile, names_to = "group") %>% 
  ggplot(aes(percentile)) +
    geom_ribbon(data = bootstrapped_cis,
                aes(ymin = lower_bound, ymax = upper_bound),
                fill = "purple", alpha = .2) +
    geom_line(aes(y = value, group = group, color = group),
              size = 2) +
    labs(x = "Percentile", y = "Value", color = NULL) +
    ggtitle("Difference to production data with bootstrapped confidence intervals") +
    scale_color_manual(values = c("black", "purple")) +
    scale_x_continuous(lim = c(0, 1), labels = scales::percent_format()) +
    theme(legend.text = element_text(size = 14),
          legend.position = "top")
```

# Time to get (more) technical

## Multiple Comparisons
```{r}
bootstrapped_quantiles
```

```{r}
# Sum of the correlations between bootstrapped differences at each quantile
cor_sum <- bootstrapped_quantiles %>%
  transmute(bootstrap, percentile, q_dif = treatment - production) %>% 
  pivot_wider(names_from = percentile, values_from = q_dif) %>% 
  select(-bootstrap) %>% 
  cor() %>% 
  .[-upper.tri(.)] %>% 
  sum()
```
```{r}
# Estimated number of independent tests
n_tests_estimate <- length(quantile_seq) ^ 2 / cor_sum
n_tests_estimate

# Bonferroni adjusted confidence interval
error_rate <- 0.05
adjusted_error_rate <- error_rate / n_tests_estimate
adjusted_error_rate
```

```{r}
1 - adjusted_error_rate
```


```{r}
bootstrapped_cis_adj <- bootstrapped_quantiles %>% 
  mutate(q_dif = treatment - production) %>% 
  group_by(percentile) %>% 
  summarise(lower_bound = quantile(q_dif, adjusted_error_rate/2),
            upper_bound = quantile(q_dif, 1 - (adjusted_error_rate / 2)))

bootstrapped_cis_adj
```

```{r ci_adjusted}
group_quantiles_df %>% 
  mutate(treatment = treatment - production,
         production = 0) %>% 
  pivot_longer(-percentile, names_to = "group") %>% 
  ggplot(aes(percentile)) +
  geom_ribbon(data = bootstrapped_cis_adj,
                aes(ymin = lower_bound, ymax = upper_bound),
                fill = "green", alpha = .2) +
    geom_ribbon(data = bootstrapped_cis,
                aes(ymin = lower_bound, ymax = upper_bound),
                fill = "purple", alpha = .2) +
    geom_line(aes(y = value, group = group, color = group),
              size = 2) +
    labs(x = "Percentile", y = "Value", color = NULL) +
    ggtitle("Difference to production data with two types of confidence intervals",
            subtitle = "Proper, Bonferroni-adjusted confidence intervals in green") +
    scale_color_manual(values = c("black", "purple")) +
    scale_x_continuous(lim = c(0, 1), labels = scales::percent_format()) +
    theme(legend.text = element_text(size = 14),
          legend.position = "top")
```

## Fast Bootstrapping on Big Data

```{r}
n_buckets <- 100

compressed_data <- tibble(production = x_production, treatment = x_treatment) %>% 
  pivot_longer(c(production, treatment), names_to = "group", values_to = "x") %>% 
  group_by(group) %>% 
  mutate(x_bucket = cut_number(x, n_buckets)) %>% 
  group_by(group, x_bucket) %>% 
  summarise(count = n(),
            x = median(x)) %>%
  ungroup()

compressed_data
```

```{r}
# Adding cumulative percentile to make search easier
compressed_data <- compressed_data %>% 
  group_by(group) %>% 
  arrange(x) %>% 
  mutate(cum_percentile = cumsum(count) / sum(count)) %>% 
  ungroup()

compressed_data
```
```{r}
# Function we can reuse to get `x` quantile from the compressed data
x_at_percentile <- function(percentile, compressed_df) {
  i <- 1
  while (compressed_df$cum_percentile[i] < percentile) {
    i <- i + 1
  }
  return (compressed_df$x[i])
}
```
```{r}
compressed_data %>% 
  filter(group == "production") %>% 
  x_at_percentile(.5, .)
```


```{r}
#' Try small comparison of approaches for `production` data
quantile_df <- tibble(percentile = seq(0, 1, by = .01)) %>%
  mutate(original = quantile(x_production, percentile)) %>%
  mutate(compressed = map_dbl(
    percentile,
    x_at_percentile,
    filter(compressed_data, group == "production")
  ))
```
```{r original_v_compressed}
quantile_df %>% 
  pivot_longer(-percentile) %>% 
  ggplot(aes(percentile, value)) +
    geom_line(aes(group = name, color = name), size = 2, alpha = .8) +
    labs(x = "Percentile", y = "Value", color = NULL) +
    ggtitle("Quantiles from original (black) and compressed (green) data") +
    scale_color_manual(values = c("green", "black")) +
    scale_x_continuous(lim = c(0, 1), labels = scales::percent_format()) +
    theme(legend.text = element_text(size = 14),
          legend.position = "top")
```

```{r}
compressed_data %>% 
  filter(group == "production") %>% 
  sample_n(n(), replace = TRUE) %>% 
  arrange(x) %>% 
  mutate(cum_percentile = cumsum(count) / sum(count))
```


# Putting it all together
